Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-13 paper/article.tex
Paper progress

\subsubsection{Simulating recovery prior to co-processor use}
We sought to additonally re-train the mRNN after applying the lesion in
order to simulate stroke recovery. For simulated lesions which zero the outputs
of neurons, the mRNN is, unfortunately, not an appropriate model.
We know that strokes can cause damage from which the subject, in many cases,
cannot recover. The mRNN model has sufficient redundancy built into it that
lesioning the outputs of large numbers of neurons leaves enough remaining
degrees of freedom that a nearly full recovery can occur, unless so many
neurons are lesioned that no stimulation could be effective.

We can, however, simulate pre-recovery in the case of a lesion that excludes
communication between the F5 and M1 modules. In that case, object shape information
cannot propagate forward in the network, in order to condition the hand for
grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp, at which
point the co-processor's job will be to forward-propagate that information.
We explore this in one of our experiments, explained below.

@@

TODO: label lines/colors

\begin{table}[]
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
Pre-recov*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
\end{tabular}
\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
\end{table}

@@

\subsection{Experiment 4: M1 lesion with pre-recovery}
In this case, we simulate some amount of recovery prior to the co-processor

>>>>

\subsubsection{Simulating recovery prior to co-processor use}
We sought to additonally re-train the mRNN after applying the lesion in
order to simulate stroke recovery. For our simulated lesions which zero
the outputs of neurons, the mRNN cannot supply an appropriate pre-recovery model.
We know that strokes can cause damage from which the subject, in many cases,
cannot recover. The mRNN model has sufficient redundancy built into it that
lesioning the outputs of large numbers of neurons leaves enough remaining
degrees of freedom that a nearly full recovery can occur, unless so many
neurons are lesioned that no stimulation could be effective.

We can, however, simulate pre-recovery in the case of a lesion that prevents
communication between the F5 and M1 modules. In that case, object shape information
cannot propagate forward in the network, in order to condition the hand for
grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp, at which
point the co-processor's job will be to forward-propagate that information.
We explore this in one of our experiments, explained below.

@@

\begin{table}[]
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
\end{tabular}
\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
\end{table}

@@

\subsection{Experiment 4: Connectivity lesion with pre-recovery}
In this case, we simulate some amount of recovery prior to the co-processor
learning. Complete recovery is impossible, because the output module's
connections to F5 (both forward and backward) are masked. As a result,
the pre-recovery causes the simulated brain to learn an object-agnostic
stereotyped grasp. The co-processor's job is to forward-propagate the
information necessary to condition the grasp on the object shape.

The co-processor in this scenario successfully improved task performance.
Note that the object separation class separation metric $S$ returned to
a healthy value towards the end of the training, indicating that the
learned solution did not simply refine the stereotyped grasp. That in
turn demonstrates the co-processor's ability to propagate information
``around'' the simulation lesion.
